### 2018-06-01 paper.tex
add writeup

\documentclass{article}
\usepackage[utf8]{inputenc}

%---------------------------------------------------
%
%---------------------------------------------------

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{blindtext}
\usepackage{bm}
\usepackage{float}
\usepackage[margin=.75in]{geometry}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage{multicol}

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\renewenvironment{abstract}
{
\begin{changemargin}{0.5in}{0.5in}
}
{
\end{changemargin}
}

\newcounter{theorems}
\newenvironment{theorem}
{
\stepcounter{theorems}
\begin{changemargin}{0.25in}{0.25in}
\noindent
\textbf{Theorem \arabic{theorems}:}
\begin{em}
}
{
\end{em}
\end{changemargin}
}

\newenvironment{proof}
{
\begin{changemargin}{0.25in}{0.25in}
\noindent
\textbf{Proof: }
}
{
\begin{flushright}
$\blacksquare$
\end{flushright}
\end{changemargin}
}

%---------------------------------------------------
%
%---------------------------------------------------

\title{Run AwAI: StarCraftII Adversary Avoidance Using Neural Networks and
  Genetic Algorithms}

\author{Kyle R. Chickering \\ krchicke@ucdavis.edu
  \and Alex R. Mirov \\ armirov@ucdavis.edu
  \and Simon H. Wu \\ simwu@ucdavis.edu
  \and Xin Jin \\ jin@ucdavis.edu}

\date{June 8, 2018}

%% TODO: Why our approach vs other approaches
%% TODO: Random action is a fairly robust approach to the problem space
%% TODO: Include convergence Graphs in the results section
%% TODO: Talk about the mutation rates in the genetic algorithm

\begin{document}

\maketitle
\hline
\\~\\

\begin{abstract}
  \textbf{Abstract}: \blindtext
  \\~\\
  \textbf{Keywords}: Machine Learning, Genetic Algorithms, Neural Networks,
  StarCraftII
\end{abstract}
\\~\\
\hline
\\~\\

\begin{multicols}{2}
\section{Introduction}
Since AlphaGo's defeat of the Go champion Ke Jie, the artificial intelligence community has
been looking for the next problem domain to apply learning techniques to. A
potentially surprising candidate for artificial intelligence testing has
emerged in the form of real time strategy games. Real time strategy games work
by having a player make critical decisions in real time against an enemy
player. The game StarCraftII, released by Blizzard entertainment in 2010, is a
real time strategy
game that has captured the interest of the artificial intelligence community.
The game presents several challenges for artificial intelligence, the biggest of
which is the magnitude of possible game states.

We present a method of teaching an agent to avoid enemy players for as long as
possibly by using genetic algorithms to evolve neural networks. We interface
our neural network agents using the PySC2 library  and evolve an agent that
shows a measurable increase in its ability to run away.

\section{Background}
This project combines two proven artificial intelligence paradigms, neural
networks and genetic algorithms. These methods are traditionally orthogonal,
but there has been a growing body of work in recent years with regards to
evolving neural networks using genetic algorithms. Most notably \cite{NEAT} and
[ref].

\subsection{Neural Networks}
Neural networks are a method of computing any computable function by using a
network of nodes called \textbf{neurons} connected by a series of edges. This
representation is inspired by the human brain, and has been proven over and over
as a heavy hitter in the world of machine learning. These networks are typically
differentiable, and by calculating error, they can be trained using a method
called backpropogation that changes the strength of the connections between
neurons.

There are two main ways to represent neural networks in software, implicitly and
explicitly. Implicit neural networks use linear algebra and a consistent layered
structure to propagate inputs forward (called feeding forward). With numerical
linear algebra libraries available for almost every major programming language,
this method of representing neural networks allows for incredibly fast
processing
of neural networks. However, this representation's greatest strength, its speed,
is the source of its greatest weakness. Implicitly defined neural networks
suffer
from the inflexibility of their layered structure. In the human brain, neurons
are not limited to any specific structure, and are free to make connections
that could not be possible in a layered model.

This leads us to explicitly defined neural networks. These networks allow the
specification of new nodes that are disconnected from the traditional layered
model. While this can be advantageous, it can also present problems for the
implementer. Because these models are represented by graphs, we can no longer
use linear algebra libraries to feed forward through the network. This causes
an increase (often significant) in running time. For small networks this is
not an issue, but for larger network this becomes (sometimes prohibitively)
problematic.

\subsection{Genetic Algorithms}
In a similar line of though that led researchers to build neural networks by
looking to nature for inspiration, genetic algorithms exploit natural selection
to create optimal solutions to problems. Research into genetic algorithms
started when researchers realized that there is nothing inherent in evolution
that limits the process to nature . In fact, by thinking of evolution as
an algorithm in and of itself, we can extend the principle of evolution to
digital systems. This is commonly done by defining what an "individual" and
"fitness function" mean in the digital evolution landscape. The algorithm
proceeds as evolution does in nature by picking the most well adapted
individuals to "breed", and evaluating their "offspring" on the same problem.
This technique has yielded several fascinating results, and often the
algorithm generates novel solutions to difficult problems.

\section{Methodology}
Our approach involved combining genetic algorithms and neural networks to
create an agent that learns through evolution how to evade antagonistic
agents. We chose this method as the problem space closely resembles nature's
predator/prey motif. By modeling our agent as an evolving neural network we can
get results that mimic generational evolution in nature. While evasion,
especially in the StarCraftII space, is a fairly simple behavior, we chose this
method as it is easily extendable to larger problem spaces by extending the
input and output dimensions.

There are many techniques to accomplish running away from enemies in
StarCraftII, for example, a simple expert system would be quite good at enemy
avoidance, however we were concerned with the scalability of our approach. We
chose the neural network and genetic algorithm approach because for dealing
with more complex tasks, our approach is more robust than a simple expert
system. Additionally, we could train various neural networks on different
tasks (avoidance, attacking, resource mining, etc.) and splice the neural
networks together to build an AI capable of multitasking during the game.

By introducing scalability, we could use our existing framework to train neural
networks in more and more difficult simulated environments, eventually leading
to an AI that could play the entire game outside of sandbox mode. Neural
networks generalize very well to a wide array of situations that are similar or
close to the situations they were trained on, and they are a perfect tool for
attacking a game that requires real time decision making and reasoning over a
complex set of inputs.

\subsection{Simulations}
We run the neural networks through a simulation to compute their fitness. The
simulations run through the PySCII interface, which runs our agent in a
simulation against pre-built AI.

We let the neural network choose the next action from our action space, which
consists of a series of movement directions (eight cardinals and a stay
command).

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{chart}
\begin{changemargin}{0.15in}{0.15in}
  \caption{Control and data flow through the Run AwAI controller.}
\end{changemargin}
\end{figure}

Talk about the mathematical model we are using to run the simulations

\subsection{Neural Network}
We started by using an explicit neural network implementation, which we chose
because we were curious if the flexible topologies would lead to a more
efficient neural network. We were inspired by the NEAT algorithm , which has
shown promise in other video game domains. However we choose a simpler
topologically explicit implementation to simplify the development of our
algorithm. We ultimately settled on an acyclic graph representation, which
allowed us the freedom to add nodes and allowed simple manipulation of
connection weights.

We chose to use our neural network to represent a function taking an input
vector consisting of relative agent position and adversary's relative position
and return an output vector of directions.
\begin{align}
  \bm{V} = \mathcal{N}(\bm{I})
\end{align}
We choose the next action according to the equation:
\begin{align}
  a_n = \max_{v \in \bm{V}} v
\end{align}
This input vector is flexible, and allows us to tweak the information that we
feed to the neural network. Additionally, while the output space we used
contains only movement in the cardinal directions, we could easily add other
supported actions like attacking and mining resources.

Our network uses the standard sigmoid function $\varsigma$ as its activation
function:
\begin{align*}
  \varsigma(x) = \frac{1}{1 + e^{-x}}
\end{align*}
We chose this function for a number of reasons, the least of which is its
popularity as an activation function in neural networks. Because of the shape
of the sigmoid function, inputs to subsequent neurons stay in the nice range of
$[-1, 1]$, which has the effect of normalizing the network. Since we are not
using backpropogation, the nice properties of the sigmoid function were
unimportant to us. However, if we were to do live backpropogation during
simulations, the sigmoid function would be a useful activation function.

When we initialize the network, we populate all of the connection weights from
a random Gaussian distribution with $\sigma = 1$ and $\mu = 0$. This serves to
make our input data ``nice'' for the sigmoid function, which ends up returning
either $1$ or $-1$ when the input data is amplified by non-gaussian data. This
also ensures that most connections are ``off'' or close to off at the start,
which allows the network to learn without any random pre-bias. Our first
generation will have very weak suggestions on the effectiveness of the next
move. This may seem bad, but it allows us to do much slower mutation changes to
the connection weights, which in turn increases the explorative power of our
algorithm. If the weights are too heavy in the beginning, the agent will favor
a local maximum almost immediately, but with a Gaussian distribution, the
algorithm explores more options before settling on a specific maximum.

\subsection{Genetic Algorithm}
For our implementation, we used a tournament style selection process to choose
our most fit individuals. We use tournament selection over other selection
methods (like probabilisitic selection), to improve convergence speeds.
Tournament selection is an exploitive selection method, and we choose exploitive
behavior over exploratory behavior because our simulation was a substantial
bottleneck in our iterative process. Because we choose an exploitive method, we
limit the chances of finding a global maximum, but increase the chances of
finding some local maximum, and finding it relatively quickly. This gave us a
chance to better observe the convergence of our algorithm, and allowed us to
demonstrate clear forward progress.

To get convergence, we spent a lot of time looking at fitness functions. We
considered many different
possibilities, but finally settled on a measure of the number of steps that our
agent took during the simulation. We figured that this would be a good measure
of our agent's ability to stay alive, and we thought that time alive was a good
indicator of the agent's ability to accomplish the goal of running away. We
computed the fitness $F$ using the following formula:
\begin{align}
  F(\bm{i}) = \mu - \eta\sqrt{\frac{1}{|\bm{\mathcal{E}}|}
  \sum_{e \in \bm{\mathcal{E}}}\left(\mathcal{S}(e) - \mu\right)^2} 
\end{align}
where
\begin{align}
  \mu = \frac{1}{|\bm{\mathcal{E}}|} \sum_{e \in \bm{\mathcal{E}}} \mathcal{S}(e)
\end{align}
and $\bm{\mathcal{E}}$ is the set of all episodes run by the PySCII simulation,
$\mathcal{S}(e)$ is the number of steps taken by the agent in the episode, and
$\bm{i}$ is a an individual of the population. $\eta$ is a hyper-parameter we
use
to fine tune the convergence of our network. We are effectively calculating the
average number of steps and penalizing the individuals based on the standard
deviation from the mean of the episodes. This serves to mitigate the effects
presented by individuals that get ``lucky'' and have a good trial runs that
offsets their average.

We seed the evolutionary controller using $n$ random individuals, and in each
evolutionary iteration, we breed $n\text{C}2$ individuals to compose the
subsequent generation. After breeding, we randomly mutate some of the children,
and then calculate the fitness for each individual in the population. To get
the next generation, we choose the most fit $n$ individuals from the population
to breed, and repeat the cycle. In addition, we found that the algorithm
converged more rapidly if we let the most fit individuals survive to the next
generation, and after making this change we noticed a jump in convergence
speeds.

The genetic algorithm relies on several hyper-parameters. The most obvious is
the
population size. We find, unsurprisingly, that larger population sizes are more
robust and generate faster convergence, this makes sense because with larger
population sizes there is an increased chance of mutation, and mutations are
largely responsible for exploratory behavior in learning algorithms. We have
three mutation parameters, which are the rate at which nodes are added to the
network, the rate at which new connections are added to the network, and the
rate at which connection weights are changed.

We mutate connection rates based on a Gaussian distribution, which, combined
with our Gaussian weight initialization, ensures that weights never change
dramatically. We adjust weights according to the formula:
\begin{align}
  w_i \leftarrow w_i + \bm{\sigma}
\end{align}
where $\bm{\sigma}$ is a random variable from a Gaussian distribution.

\subsection{Controller}

\subsection{PySC2 Agent Interface}

\section{Results}

\section{Conclusion}
Different selection types?
\subsection{Improvements}
\subsection{Future Work}

\end{multicols}

\\~\\
\hline
\\~\\

\begin{thebibliography}{9}
\bibitem{NEAT} 
  Kenneth O. Stanley, Risto Miikkulainen.
  \textit{Evolving Neural Networks through Augmenting Topologies}. 
  The MIT Press Journals.

\bibitem{dejong}
  Kenneth A. De Jong. \textit{Evolutionary Computation: A Unified Approach}.
  MIT Press, Feb 3, 2006.

\bibitem{russell}
  Stuart Russell, Peter Norvig.
  \textit{Artificial Intelligence: A Modern Approach}.
  Upper Saddle River (New Jersey, 1995).

\bibitem{lehman}
  Joel Lehman et al.
  \textit{The Surprising Creativity of Digital Evolution: A Collection of
    Anecdotes from the Evolutionary Computation and Artificial Life Research
    Communities}.
  \textbf{https://arxiv.org/pdf/1803.03453.pdf}.

\bibitem{pysc2}
  Google DeepMind pysc2.
  \textbf{https://github.com/deepmind/pysc2}.

\end{thebibliography}

\end{document}

### 2018-06-01 paper.tex
added blurb on pysc2 interface
Added a section on map variations and learning environments

@@
 \centering
-\includegraphics[width=0.45\textwidth]{chart}
+%\includegraphics[width=0.45\textwidth]{chart}
 \begin{changemargin}{0.15in}{0.15in}
@@
 
-\subsection{PySC2 Agent Interface}
+\subsection{PySC2 Interface}
+
+The learning environment was based on the existing mini game environments from the PySC2 API. Specifically, our learning environment was built on top of the DefeatRoaches mini game.
+
+The maps were retrieved from Deepmind's PySC2 Github repository and 
+put into the StarcraftII/Maps/ directory.
+
+All maps were edited in the SCII map editor.
+
+\subsubsection{Learning Environment}
+The DefeatRoaches mini game is 64x64 simple map without any obstacles that spawns a group of marines and roaches at two spawn points within the map.
+
+\subsubsection{Map Variations}
+DefeatRoaches 1v1 - 
+Similar to the DefeatRoaches map except for the number of marines and 
+roaches in the map at spawn time. This custom map only has 1 marine and 1 roach at the designated spawn points.
+
+The decision to make this custom map was made early on to make the learning environment as simple as possible for training purposes.
+
+DefeatRoaches96 - 
+This variation, DefeatRoaches96, is a 96x96 map that is bigger than the original. This allows the agent to have a larger input space to choose from which could potentially improve rate of convergence.
 

### 2018-06-02 paper.tex
more papering

--- 2018-06-01 paper.tex
@@
 \section{Introduction}
-Since AlphaGo's defeat of the Go champion Ke Jie, the artificial intelligence community has
+Since AlphaGo's defeat of the Go champion Ke Jie, the artificial intelligence
+community has
 been looking for the next problem domain to apply learning techniques to. A
@@
 \end{align}
-We choose the next action according to the equation:
+Where $\mathcal{N}$ is the function computed by our neural network. We choose
+the next action according to the equation:
 \begin{align}
-  a_n = \max_{v \in \bm{V}} v
+  a_n = \max_{v \in \bm{V}} \: v
 \end{align}
@@
 \section{Results}
+%% TODO: Talk about choice of hyper paramters, population size, episode count
+%% mutation constants, eta in fitness function
 
@@
 Different selection types?
+
 \subsection{Improvements}
+
 \subsection{Future Work}
+This project can certainly be extended in many ways, for example, we could
+extend the output vectors to include offinsive manuvering by our agent, and
+train the agent to be agressive by considering the numebr of adversaries killed
+by our agent during the simulation.
 

### 2018-06-04 paper.tex
update paper, add figures

@@
 %% TODO: Talk about the mutation rates in the genetic algorithm
+%% TODO: Reference the fact that neuro evolution is good in high dimensional
+%%    and continuous state spaces (ref neat paper)
 
@@
 \subsection{Simulations}
-We run the neural networks through a simulation to compute their fitness. The
-simulations run through the PySCII interface, which runs our agent in a
-simulation against pre-built AI.
+To compute the fitness for each neural network in the population, we run the
+network through a simulation. The simulation is run in StarCraftII, and our
+simulation controller is built on top of the PySCII interface, and simulates
+a ``game'' with our agent and a group of adversaries, which are controlled by
+the ingame AI that come with the game.
+
+Each simulation is called an episode, and we ran the networks through 5-10
+episodes each so that we could get a representative sample of the amount of time
+that each nerual network was able to survive.
 
@@
 \centering
-%\includegraphics[width=0.45\textwidth]{chart}
+\includegraphics[width=0.45\textwidth]{chart}
 \begin{changemargin}{0.15in}{0.15in}
@@
 because we were curious if the flexible topologies would lead to a more
-efficient neural network. We were inspired by the NEAT algorithm , which has
+efficient neural network. We were inspired by the NEAT algorithm \cite{NEAT}, which has
 shown promise in other video game domains. However we choose a simpler
@@
 computed the fitness $F$ using the following formula:
-\begin{align}
-  F(\bm{i}) = \mu - \eta\sqrt{\frac{1}{|\bm{\mathcal{E}}|}
-  \sum_{e \in \bm{\mathcal{E}}}\left(\mathcal{S}(e) - \mu\right)^2} 
-\end{align}
-where
-\begin{align}
-  \mu = \frac{1}{|\bm{\mathcal{E}}|} \sum_{e \in \bm{\mathcal{E}}} \mathcal{S}(e)
+% \begin{align}
+%   F(\bm{i}) = \mu - \eta\sqrt{\frac{1}{|\bm{\mathcal{E}}|}
+%   \sum_{e \in \bm{\mathcal{E}}}\left(\mathcal{S}(e) - \mu\right)^2} 
+% \end{align}
+% where
+% \begin{align}
+%   \mu = \frac{1}{|\bm{\mathcal{E}}|} \sum_{e \in \bm{\mathcal{E}}} \mathcal{S}(e)
+    %   \end{align}
+\begin{align}\label{eq:fitness}
+  F(\bm{i}) = \frac{1}{|\bm{\mathcal{E}}|} \sum_{e \in \bm{\mathcal{E}}} \mathcal{S}(e) - \eta\sqrt{\frac{1}{|\bm{\mathcal{E}}|}
+   \sum_{e \in \bm{\mathcal{E}}}\left(\mathcal{S}(e) - \frac{1}{|\bm{\mathcal{E}}|} \sum_{e \in \bm{\mathcal{E}}} \mathcal{S}(e)\right)^2} 
 \end{align}
@@
 \subsection{PySC2 Interface}
+Our simulation learning environment was based on an existing environment
+included in the PySC2 library\cite{pysc2}. We use the ``Defeat Roaches'' mini
+game as the baseline for our learning. Once we had the map, we used the
+StarCraftII map editor to tweak the mini game to better suit our needs. We
+created three maps that were used in testing. The defualt map is a simple square
+map which spawns a group of adversarialy ``roachs'', and a group of ``marines'',
+who are controlled by our agent. The marines and roaches are spawned in the
+same place every time, which gives us a consistent platform to test our agent.
+
+The second variation is similar to the first, except that there is only one
+``roach'' and one ``marine''. This map type highlights the capabilities of our
+agent in a more intimate setting. The group of marines has the advantage of
+being able to get ``lucky'' and have some members survive, but the one on one
+variation forces the neural network to work well from the beginning.
 
@@
 \section{Results}
+We ran several simulations with various choices of hyperparameters, and our
+results are promising. We saw measurable improvement in preformence after just
+a few iterations of the algorithm, which gave us a steep initial learning curve
+before the improvements started to taper off.
+
+\begin{figure}[H]\label{fig:fitness graph}
+\centering
+\includegraphics[width=0.45\textwidth]{fig_1}
+\begin{changemargin}{0.15in}{0.15in}
+  \caption{Graph of average fitness (blue) and top fitness (orange) for each
+    generation in Trial 1. Top fitness is the fitness of the most fit individual
+    at the end of each generation's simulations. Average fitness is calculated
+    by averaging \eqref{eq:fitness} over all individuals $\bm{i}$ in the
+    population. The breeding constant is set to 15, node generation rate is set
+    to 0.3, new connections at 0.2 and non structural mutations at a rate of
+    0.4}
+\end{changemargin}
+\end{figure}
+
+We see in Figure \ref{fig:fitness graph} an artifact from our choice of genetic
+algorithm. Because we keep the top individual from each generation in the gene
+pool, we see that the top individual has palteaued early, and no new individual
+is achieving a higer fitness. We see that the average is countinually decreasing
+though, which will increase the chances that an individual will be bred that has
+a higher fitness than the individual that is dominating the maximum fitness.
+
+We can also see that the average sometimes plateaus as well. This is due to a
+homonginization of the gene pool, where the children individuals are very
+similar to their parents. This is a case of finding a local maximum. Since we
+favored an exploitive algorithm, this is to be expected, however we see that the
+population can still achieve higher fitness meausures which means that the
+algorithm still retains some amount of exploratory behavior.
+
+\begin{figure}[H]\label{fig:fitness graph}
+\centering
+\includegraphics[width=0.45\textwidth]{fig_2}
+\begin{changemargin}{0.15in}{0.15in}
+  \caption{Graph of average fitness (blue) and top fitness (orange) for each
+    generation in Trial 1. Top fitness is the fitness of the most fit individual
+    at the end of each generation's simulations. Average fitness is calculated
+    by averaging \eqref{eq:fitness} over all individuals $\bm{i}$ in the
+    population. The breeding constant is set to 15, node generation rate is set
+    to 0.5, new connections at 0.5 and non structural mutations at a rate of
+    0.75}
+\end{changemargin}
+\end{figure}
+
+In our second trial, we see that the average fitness increases much more
+regularly. However, despite having more generations, trial two did not reach the
+same average fitness as trial one did. One possible reason for this is since our
+mutation rates are higher, our algorithm favored a more exploratory approach,
+which lead to a more consistant incresase in fitness, but did not reach as high
+of an average fitness.
+
 %% TODO: Talk about choice of hyper paramters, population size, episode count
@@
 Different selection types?
+\blindtext
 
 \subsection{Improvements}
+\blindtext
+\blindtext
 
@@
 extend the output vectors to include offinsive manuvering by our agent, and
-train the agent to be agressive by considering the numebr of adversaries killed
+train the agent to be agressive by considering the number of adversaries killed
 by our agent during the simulation.
+
+Another possibility that we did not explore, was the effect of removing neurons
+from the network. Neural networks can often be plauged by overfitting input
+data, and by removing neurons we hypothesize that any overfitting we encounter
+could be mitigated. We would remove these neurons in the same way that we add
+neurons, that is, non-deterministically as a possible mutation. In the study of
+biological creatures, we often see either total dropping of traits from the
+population genome or the transition to vestigial genetic structure
+\cite{lehman}. In keeping with this, instead of outright removing the neuron, we
+could significantly decrease the strength of the connections to it, essentially
+removing it from the network. If the neuron in question was indeed problematic,
+we would see it countinue to evolve very low weights, essentially pruning the
+neuron and rendering it vestigial.
+
+We could also track the lifespan of certain individuals and kill them off after
+a certain number of generations. We found that the algorithm would sometimes
+plateau with a very fit individual for a period of time, and it would be
+interesting to see what effects a notion of ``lifetime'' would have on the
+algorithm as a whole.
 

### 2018-06-04 doc/run_awai.tex
refined pysc2 interfaces section

--- 2018-06-04 paper.tex
@@
 created three maps that were used in testing. The defualt map is a simple square
-map which spawns a group of adversarialy ``roachs'', and a group of ``marines'',
+map which spawns a group of adversarialy ``roaches'', and a group of ``marines'',
 who are controlled by our agent. The marines and roaches are spawned in the
@@
 agent in a more intimate setting. The group of marines has the advantage of
-being able to get ``lucky'' and have some members survive, but the one on one
+being able to get ``lucky'' and have some members survive, but the 1v1
 variation forces the neural network to work well from the beginning.
-
-The learning environment was based on the existing mini game environments from the PySC2 API. Specifically, our learning environment was built on top of the DefeatRoaches mini game.
-
-The maps were retrieved from Deepmind's PySC2 Github repository and 
-put into the StarcraftII/Maps/ directory.
 
@@
 \subsubsection{Learning Environment}
-The DefeatRoaches mini game is 64x64 simple map without any obstacles that spawns a group of marines and roaches at two spawn points within the map.
+The final map chosen for our simulations was the default DefeatRoaches mini game.
+The mini game provides a simple map that is 64x64 without any obstacles that spawns a group of marines and roaches at two spawn points within the map.
 
@@
 DefeatRoaches 1v1 - 
-Similar to the DefeatRoaches map except for the number of marines and 
+Similar to the default DefeatRoaches map except for the number of marines and 
 roaches in the map at spawn time. This custom map only has 1 marine and 1 roach at the designated spawn points.
@@
 DefeatRoaches96 - 
-This variation, DefeatRoaches96, is a 96x96 map that is bigger than the original. This allows the agent to have a larger input space to choose from which could potentially improve rate of convergence.
+This variation, DefeatRoaches96, is a 96x96 map that is bigger than the original.
+This allows the agent to have a larger input space to choose from which could potentially improve rate of convergence to an optimal solution.
 

### 2018-06-04 doc/run_awai.tex
fixed minor typos

@@
 We present a method of teaching an agent to avoid enemy players for as long as
-possibly by using genetic algorithms to evolve neural networks. We interface
+possible by using genetic algorithms to evolve neural networks. We interface
 our neural network agents using the PySC2 library  and evolve an agent that
@@
 \section{Background}
-This project combines two proven artificial intelligence paradigms, neural
+This project combines two proven artificial intelligence paradigms: neural
 networks and genetic algorithms. These methods are traditionally orthogonal,
@@
 a ``game'' with our agent and a group of adversaries, which are controlled by
-the ingame AI that come with the game.
+the in game AI that come with the game.
 
@@
 episodes each so that we could get a representative sample of the amount of time
-that each nerual network was able to survive.
+that each neural network was able to survive.
 
@@
 our most fit individuals. We use tournament selection over other selection
-methods (like probabilisitic selection), to improve convergence speeds.
+methods (like probabilistic selection), to improve convergence speeds.
 Tournament selection is an exploitive selection method, and we choose exploitive

### 2018-06-04 doc/run_awai.tex
added controller writeup

--- 2018-06-04 doc/run_awai.tex
@@
 
+The Pysc2 interface allowed us to create an agent to interact with the Star
+Craft II game.  The agent class contains a \texttt{step()} function that is called
+several times per second of game time. This meant that all of the actions the
+agent took had to be in this step function, and thus this is where we would
+integrate our neural net.  We decided to use the mini game map DefeatRoaches,
+which is included with Pysc2 as a base for our simulation environment. By
+heavily relying on a preexisting map and mini game, we could focus more of our
+time and resources on developing the genetic algorithm and neural network
+implementation.
+
+To make our controller implementation as modular as possible, we decided to
+provide a dictionary of inputs and receive a dictionary of outputs. The inputs
+would be the x and y coordinates of the agent’s units, as well as the x and y
+coordinates of the enemy units. The inputs had to be created and assembled into
+a dictionary in the step function.  In order to normalize the inputs, we divided
+the coordinate value by the maximum map size in the corresponding axis.  To
+obtain the values for the inputs, we created a series of methods for our agent
+class.  These methods include \texttt{getCurrentLocation()} and
+\texttt{getCurrentEnemeyLocation}().  We also created methods to move the friendly units,
+including \texttt{moveToLocation()} and setTargetDestination.  These functions could be
+used together to set a target destination coordinate and cause the friendly
+units to move toward that target.
+
+One problem we encountered that we had to correct was determining when units
+arrived at their target destination.  We solved this by comparing the target
+destination and the units current location every time \texttt{step()} is called.  Because
+the DefeatRoaches mini game uses multiple units (both friendly and hostile), the
+unit location coordinates are averages of all selected units locations.  This
+means that often units will never actually arrive at their precise destination,
+just get very close to it.  To mitigate this issue, we decided to consider units
+within a certain radius of their target destination as at their destination.
+Once this condition is met, the \texttt{moveToLocation()} function returns a status of
+arrived at target.
+
+Whenever \texttt{step()} is called by Pysc2, the first thing we do is check the
+observable action space. If friendly unit movement is not an option, this means
+the unit(s) are not selected.  If units are not selected, this means it is the
+first time the \texttt{step()} function is being called in the simulation episode.  In
+this case, in addition to selecting the units, we also reset various counters
+(such as steps survived).  During the first call of \texttt{step()}, we also unpickle a
+neural network file.
+
+In the case where \texttt{step()} is called and units are already selected, the
+\texttt{moveToLocation()} function is called.  Initially, we set the target destination
+to the enemy location.  We have our units move directly to the hostile units in
+order to engage them in combat, and this initiate a pursuit from the enemy.  We
+felt this would be the most direct way to initiate a pursuit.  We investigated
+running two Pysc2 agents against each other, however Pysc2 currently only
+support the simulation of one custom agent at a time. Once the friendly units
+arrive at their destination, the \texttt{moveToLocation()} function returns an arrived at
+target status to the calling \texttt{step()} function.
+
+When the \texttt{step()} function receives an arrived at target status, a new target
+destination is set.  To generate a new target destination, the relevant aspects
+of the current state (current location of friendly and enemy units) are captured
+and fed into the neural network.  The neural network returns a dictionary
+containing the action space and a value between -1 and 1 for each action label
+key.  The action with the highest number will be chosen, and then taken.
+
+Our action space is movements that the friendly units can take.  We decided to
+quantize the action space into discrete movements, as opposed to a continuous
+x-y coordinate plane. By only allowing our units to move in the 4 cardinal
+directions, diagonals, or stay put, we hoped our first implementation of the
+neural network and genetic algorithm would have an easier time converging.  In
+the future, we could expand the output action space to a continuous x and y
+coordinate plane, as opposed to discrete directions. Although a larger action
+space can often require more iterations of the genetic algorithm to converge, as
+the problem complexity is much higher.
+
+To evaluate the fitness of the neural network, we count the number of steps that
+the units survived for.  One problem we encountered and solved, was the initial
+series of steps taken before engaging the enemy units. For example, the
+DefeatRoaches map spawns the enemy and friendly units on opposite sides of the
+map, in a random selection of several predefined start locations.  However, the
+distance between these start locations can vary between simulation episode.
+This means our neural network fitness (number of steps survived) could be
+slightly affected by an irrelevant condition.  To mitigate this issue, we
+subtract the number of steps it takes until the friendly units reach their first
+destination (the location of the enemy units) and start their engagement. By
+only counting steps after enemy engagement has begun, our fitness function is
+more reflective of survival time.
+
+In order for our evolutionary genetic algorithm to evolve, the end fitness of a
+network must be returned from the Pysc2 simulation.  Our simulation runs a
+configurable (up to 55) episodes per individual (neural network) and averages
+the fitness for all of those simulation episodes. We played with the number of
+simulation episodes per simulation test, exploring the tradeoff between
+simulation runtime and precision.  In the future, we could explore modifying our
+fitness function to incorporate more meaning from having multiple simulation
+episodes.  One future idea we had was to see if the group of individual survival
+times were skewed above or below the mean.
+
+The average fitness for the individual is then saved in a pickle file, along
+with the fitness of the top (most fit) individual.  Because the Pysc2 simulation
+is invoked from the genetic algorithm code, the algorithm can then unpickle the
+file.  After all individuals in a generation are tested and their fitnesses
+collected, the genetic algorithm can select the most fit individuals (neural
+networks) for breeding.
+
 \subsection{PySC2 Interface}

### 2018-06-04 pysc2/agents/RunAwAI_Agent.py
misc minor cleanup

@@ -2,11 +2,11 @@
 # Simon Wu for ECS 170 @ UC Davis, Spring 2018.
 
 """
-  RunAwAI Agent
-  Will run away from enemy units.
+RunAwAI Agent
+Will run away from enemy units.
 
-  TEST USAGE:
-    python -m pysc2.bin.agent --map DefeatRoaches --agent pysc2.agents.RunAwAI_Agent.RunAwAI
+TEST USAGE:
+  python -m pysc2.bin.agent --map DefeatRoaches --agent pysc2.agents.RunAwAI_Agent.RunAwAI
 """
 
 from __future__ import absolute_import
@@ -62,7 +62,7 @@ class RunAwAI(base_agent.BaseAgent):
 
   def getCurrentLocation(self):
     """
-      returns [xcoord, ycoord] of current player location
+    returns [xcoord, ycoord] of current player location
     """
     player_relative = self.obs.observation["screen"][_PLAYER_RELATIVE]
     player_y, player_x = (player_relative == _PLAYER_FRIENDLY).nonzero()
@@ -70,7 +70,7 @@ class RunAwAI(base_agent.BaseAgent):
 
   def getCurrentEnemyLocation(self):
     """
-      returns [xcoord, ycoord] of current enemy location
+    returns [xcoord, ycoord] of current enemy location
     """
     player_relative = self.obs.observation["screen"][_PLAYER_RELATIVE]
     player_y, player_x = (player_relative == _PLAYER_HOSTILE).nonzero()
@@ -79,12 +79,12 @@ class RunAwAI(base_agent.BaseAgent):
 
   def moveToLocation (self):
     """
-      Moves selected unit(s) to self.target location. Returns an object with the
-      movement status of the units, i.e. if they're still enroute to target, or
-      if they have arrived, and the return function that actually invokes the 
-      movement. 
-      The returnObject["function"] must be returned in the main step()
-      function to invoke movement.
+    Moves selected unit(s) to self.target location. Returns an object with the
+    movement status of the units, i.e. if they're still enroute to target, or
+    if they have arrived, and the return function that actually invokes the 
+    movement. 
+    The returnObject["function"] must be returned in the main step()
+    function to invoke movement.
     """
     target = self.target
     # get current location
@@ -109,7 +109,7 @@ class RunAwAI(base_agent.BaseAgent):
 
   def setTargetDestination (self, coordinates):
     """
-      Sets the target destination to an x and y tuple in coordinates
[…]

### 2018-06-04 doc/run_awai.tex
spelling
- fixed spelling in run_awai.tex
- updated fig_2 to use same scale as fig_1

--- 2018-06-04 doc/run_awai.tex
@@
 community has
-been looking for the next problem domain to apply learning techniques to. A
-potentially surprising candidate for artificial intelligence testing has
+been looking for a new problem domain to apply learning techniques to. A
+surprising candidate for artificial intelligence testing has
 emerged in the form of real time strategy games. Real time strategy games work
@@
 The game presents several challenges for artificial intelligence, the biggest of
-which is the magnitude of possible game states.
+which is the magnitude of possible game states. The game technically has a
+finite number of game states, but the granularity of those states renders the
+game practically continuous.
 
 We present a method of teaching an agent to avoid enemy players for as long as
-possible by using genetic algorithms to evolve neural networks. We interface
-our neural network agents using the PySC2 library  and evolve an agent that
-shows a measurable increase in its ability to run away.
+possibly by using genetic algorithms to evolve neural networks (neuroevolution).
+We interface our neural network agents using the PySC2 library and evolve an
+agent that shows a measurable increase in its ability to run away from
+adversarial opponents.
 
 \section{Background}
-This project combines two proven artificial intelligence paradigms: neural
+This project combines two proven artificial intelligence paradigms, neural
 networks and genetic algorithms. These methods are traditionally orthogonal,
 but there has been a growing body of work in recent years with regards to
-evolving neural networks using genetic algorithms. Most notably \cite{NEAT} and
-[ref].
+evolving neural networks using genetic algorithms, most notably \cite{NEAT}.
 
@@
 representation is inspired by the human brain, and has been proven over and over
-as a heavy hitter in the world of machine learning. These networks are typically
+as a heavy hitter in the world of machine learning. These networks are
 differentiable, and by calculating error, they can be trained using a method
 called backpropogation that changes the strength of the connections between
-neurons.
+neurons. The backpropogation method is the status quo of training neural
+networks, however, alternative methods exist, such as neuroevolution.
+Neuroevolution is an efficient alternative, especially in spaces with
+continuous state spaces.
 
@@
 started when researchers realized that there is nothing inherent in evolution
-that limits the process to nature . In fact, by thinking of evolution as
-an algorithm in and of itself, we can extend the principle of evolution to
+that limits the process to nature \cite{dejong}. In fact, by thinking of
+evolution as an algorithm in and of itself, we can extend the principles of
+evolution to
 digital systems. This is commonly done by defining what an "individual" and
@@
 This technique has yielded several fascinating results, and often the
-algorithm generates novel solutions to difficult problems.
+algorithm generates novel solutions to difficult problems \cite{lehman}.
 
 \section{Methodology}
-Our approach involved combining genetic algorithms and neural networks to
+Our approach involves combining genetic algorithms and neural networks to
 create an agent that learns through evolution how to evade antagonistic
 agents. We chose this method as the problem space closely resembles nature's
-predator/prey motif. By modeling our agent as an evolving neural network we can
+predator/prey motif, and with such a large state space, neuroevolution is
+actually an efficient solution \cite{NEAT}. By modeling our agent as an
+evolving neural network we can
 get results that mimic generational evolution in nature. While evasion,
@@
 networks in more and more difficult simulated environments, eventually leading
-to an AI that could play the entire game outside of sandbox mode. Neural
+to an AI that could play the entire game outside of a sandbox mode. Neural
 networks generalize very well to a wide array of situations that are similar or
@@
 simulation controller is built on top of the PySCII interface, and simulates
-a ``game'' with our agent and a group of adversaries, which are controlled by
-the in game AI that come with the game.
+a ``game'' between our agent and a group of adversaries, which are controlled by
+the in-game AI that ships with the game.
 
@@
 consists of a series of movement directions (eight cardinals and a stay
-command).
+command). The game runs a ``continuous'' time simulation (obviously continuity
+is limited by the computer's capabilities), but we chose to have our neural
+network make decisions in finite time intervals. We did this so that we would
+not fall victim to our network not making decisions fast enough, and chose a
+time step that would allow the network to make decisions quickly enough to
+function well.
 
@@
 \centering
-\includegraphics[width=0.45\textwidth]{chart}
+\includegraphics[width=0.45\textwidth]{chart}\label{fig:flow chart}
 \begin{changemargin}{0.15in}{0.15in}
@@
 
-Talk about the mathematical model we are using to run the simulations
+Our whole system ran as outlined in \ref{fig:flow chart}. We decoupled the
+various aspects of the system as outlined as a good architecture practice in
+\cite{pragprog}. We run a general simulation controller, which interfaces with
+the neural networks through a module, and after creating and manipulating the
+networks, sends them to a simulation controller which interfaces with PySC2,
+which runs the simulations in StarCraftII.
+
+We collect from each simulation the set $\bm{\mathcal{E}}$ of episodes, which we
+can use to collect data from the simulation. We also compute the number of steps
+taken by our agent in each simulation, denoted
+$\mathcal{S}(e\in\bm{\mathcal{E}})$. The steps are the number of actions that
+the agent made during the simulation, and we chose to measure this because it is
+a good indicator of success of the agent at running away.
 
@@
 shown promise in other video game domains. However we choose a simpler
-topologically explicit implementation to simplify the development of our
+topologically explicit implementation to simplify and accelerate the development
+of our
 algorithm. We ultimately settled on an acyclic graph representation, which
@@
 We chose to use our neural network to represent a function taking an input
-vector consisting of relative agent position and adversary's relative position
-and return an output vector of directions.
+vector $\bm{I}$ consisting of relative agent position and adversary's relative
+position and return an output vector $\bm{V}$ of floating point values
+corresponding to the agent's belief that a particular action would lead to a
+successful outcome. The neural network is a function
+$\mathcal{N}:\text{dim}(\bm{I})\rightarrow\text{dim}(\bm{V})$:
 \begin{align}
@@
 \end{align}
-Where $\mathcal{N}$ is the function computed by our neural network. We choose
-the next action according to the equation:
+We choose the next action according to the simple maximization equation:
 \begin{align}
@@
 
-Our network uses the standard sigmoid function $\varsigma$ as its activation
-function:
+Our neural network uses the standard sigmoid function $\varsigma$ as its
+activation function:
 \begin{align*}
@@
 We chose this function for a number of reasons, the least of which is its
-popularity as an activation function in neural networks. Because of the shape
-of the sigmoid function, inputs to subsequent neurons stay in the nice range of
+popularity as an activation function in neural networks. Because of the shape of
+the sigmoid function, inputs to subsequent neurons stay in the nice range of
 $[-1, 1]$, which has the effect of normalizing the network. Since we are not
@@
 unimportant to us. However, if we were to do live backpropogation during
-simulations, the sigmoid function would be a useful activation function.
+simulations (See section \ref{sec:improvements}), the sigmoid function would be
+a useful activation function.
 
-When we initialize the network, we populate all of the connection weights from
-a random Gaussian distribution with $\sigma = 1$ and $\mu = 0$. This serves to
+When we initialize the network, we populate all of the connection weights from a
+random Gaussian distribution with $\sigma = 1$ and $\mu = 0$. This serves to
 make our input data ``nice'' for the sigmoid function, which ends up returning
@@
 generation will have very weak suggestions on the effectiveness of the next
-move. This may seem bad, but it allows us to do much slower mutation changes to
-the connection weights, which in turn increases the explorative power of our
-algorithm. If the weights are too heavy in the beginning, the agent will favor
-a local maximum almost immediately, but with a Gaussian distribution, the
-algorithm explores more options before settling on a specific maximum.
+move. This may seem counter-productive, but it allows us to do much slower
+mutation changes to the connection weights, which in turn increases the
+explorative power of our algorithm at the start of the evolutions. If the
+weights are too heavy in the beginning, the agent will favor a local maximum
+almost immediately, but with a Gaussian distribution, the algorithm explores
+more options before settling on a specific maximum.
 
@@
 our most fit individuals. We use tournament selection over other selection
-methods (like probabilistic selection), to improve convergence speeds.
+methods (like probabilisitic selection), to improve convergence speeds.
 Tournament selection is an exploitive selection method, and we choose exploitive
@@
 To get convergence, we spent a lot of time looking at fitness functions. We
-considered many different
-possibilities, but finally settled on a measure of the number of steps that our
-agent took during the simulation. We figured that this would be a good measure
-of our agent's ability to stay alive, and we thought that time alive was a good
-indicator of the agent's ability to accomplish the goal of running away. We
-computed the fitness $F$ using the following formula:
-% \begin{align}
-%   F(\bm{i}) = \mu - \eta\sqrt{\frac{1}{|\bm{\mathcal{E}}|}
-%   \sum_{e \in \bm{\mathcal{E}}}\left(\mathcal{S}(e) - \mu\right)^2} 
-% \end{align}
-% where
-% \begin{align}
-%   \mu = \frac{1}{|\bm{\mathcal{E}}|} \sum_{e \in \bm{\mathcal{E}}} \mathcal{S}(e)
-    %   \end{align}
+considered many different possibilities, but finally settled on a measure of the
+number of steps that our agent took during the simulation, with a penalty for
+inconsistency. We figured that this would be a good measure of our agent's
+ability to stay alive, and we thought that time alive was a good indicator of
+the agent's ability to accomplish the goal of running away. We computed the
+fitness $F$ using the following formula:
 \begin{align}\label{eq:fitness}
-  F(\bm{i}) = \frac{1}{|\bm{\mathcal{E}}|} \sum_{e \in \bm{\mathcal{E}}} \mathcal{S}(e) - \eta\sqrt{\frac{1}{|\bm{\mathcal{E}}|}
-   \sum_{e \in \bm{\mathcal{E}}}\left(\mathcal{S}(e) - \frac{1}{|\bm{\mathcal{E}}|} \sum_{e \in \bm{\mathcal{E}}} \mathcal{S}(e)\right)^2} 
+  F(\bm{i}) = \frac{1}{|\bm{\mathcal{E}}|} \sum_{e
+\in \bm{\mathcal{E}}} \mathcal{S}(e) - \eta\sqrt{\frac{1}{|\bm{\mathcal{E}}|}
+\sum_{e \in \bm{\mathcal{E}}}\left(\mathcal{S}(e) - \frac{1}{|\bm{\mathcal{E}}|}
+\sum_{e \in \bm{\mathcal{E}}} \mathcal{S}(e)\right)^2}
 \end{align}
-and $\bm{\mathcal{E}}$ is the set of all episodes run by the PySCII simulation,
-$\mathcal{S}(e)$ is the number of steps taken by the agent in the episode, and
-$\bm{i}$ is a an individual of the population. $\eta$ is a hyper-parameter we
-use
-to fine tune the convergence of our network. We are effectively calculating the
-average number of steps and penalizing the individuals based on the standard
-deviation from the mean of the episodes. This serves to mitigate the effects
-presented by individuals that get ``lucky'' and have a good trial runs that
-offsets their average.
+where $\bm{\mathcal{E}}$ is the set of all episodes run by the PySC2
+simulation, $\mathcal{S}(e)$ is the number of steps taken by the agent in the
+episode, and $\bm{i}$ is a an individual of the population. $\eta$ is a
+hyper-parameter we use to fine tune the convergence of our network. We are
+effectively calculating the average number of steps and penalizing the
+individuals based on the standard deviation from the mean of the episodes. This
+penalty serves to mitigate the effects presented by individuals that get
+``lucky'' and have a few good trial runs that offset their average.
 
@@
 subsequent generation. After breeding, we randomly mutate some of the children,
-and then calculate the fitness for each individual in the population. To get
-the next generation, we choose the most fit $n$ individuals from the population
-to breed, and repeat the cycle. In addition, we found that the algorithm
-converged more rapidly if we let the most fit individuals survive to the next
-generation, and after making this change we noticed a jump in convergence
-speeds.
+and then calculate the fitness for each individual in the population. To get the
+next generation, we choose the most fit $n$ individuals from the population to
+breed, and repeat the cycle. In addition, we found that the algorithm converged
+more rapidly if we let the most fit individuals survive to the next generation,
+and after making this change we noticed a jump in convergence rates.
 
 The genetic algorithm relies on several hyper-parameters. The most obvious is
-the
-population size. We find, unsurprisingly, that larger population sizes are more
-robust and generate faster convergence, this makes sense because with larger
-population sizes there is an increased chance of mutation, and mutations are
-largely responsible for exploratory behavior in learning algorithms. We have
-three mutation parameters, which are the rate at which nodes are added to the
-network, the rate at which new connections are added to the network, and the
-rate at which connection weights are changed.
+the population size. We find, unsurprisingly, that larger population sizes are
+more robust and promote faster convergence, this makes intuitive sense because
+with larger population sizes there is an increased chance of mutation, and
+mutations are largely responsible for exploratory behavior in learning
+algorithms. We have three mutation parameters, which are the rate at which nodes
+are added to the network, the rate at which new connections are added to the
+network, and the rate at which connection weights are changed.
 
@@
 \end{align}
-where $\bm{\sigma}$ is a random variable from a Gaussian distribution.
+where $\bm{\sigma}$ is a random variable from a Gaussian
+distribution. We found that changing the weights according to a Gaussian
+distribution promoted faster convergence than using weights taken from a uniform
+distribution.
 
@@
 StarCraftII map editor to tweak the mini game to better suit our needs. We
-created three maps that were used in testing. The defualt map is a simple square
-map which spawns a group of adversarialy ``roaches'', and a group of ``marines'',
-who are controlled by our agent. The marines and roaches are spawned in the
-same place every time, which gives us a consistent platform to test our agent.
+created three maps that were used in testing. The default map is a simple square
+map which spawns a group of adversarial ``roaches'', and a group of ``marines'',
+who are controlled by our agent. The marines and roaches are spawned in the same
+place every time, which gives us a consistent platform to test our agent.
 
@@
 agent in a more intimate setting. The group of marines has the advantage of
-being able to get ``lucky'' and have some members survive, but the 1v1
-variation forces the neural network to work well from the beginning.
+being able to get ``lucky'' and have some members survive, but the one-on-one
+variation forces the neural network to work well from the very beginning.  With
+the one-on-one map, we were hoping to highlight more explicitly the capabilities
+of our trained neural networks.
 
-All maps were edited in the SCII map editor.
+The third map variation was a larger sized version of the original map, which we
+made in the hopes that we would see different convergence rates and survival
+strategies on a larger game space. We were hoping to see the agent take
+advantage of the larger space, but unfortunately did not have time to run
+extensive simulations on this map.
 
-\subsubsection{Learning Environment}
-The final map chosen for our simulations was the default DefeatRoaches mini game.
-The mini game provides a simple map that is 64x64 without any obstacles that spawns a group of marines and roaches at two spawn points within the map.
+All three maps have a slight bit of non-determinism built in, which means
+subsequent simulations using the same neural network will not be identical. This
+gave us the opportunity to gather more realistic data about the performance of
+our agent, since we could accumulate an average over a number or trials.
 
-\subsubsection{Map Variations}
-DefeatRoaches 1v1 - 
-Similar to the default DefeatRoaches map except for the number of marines and 
-roaches in the map at spawn time. This custom map only has 1 marine and 1 roach at the designated spawn points.
-
-The decision to make this custom map was made early on to make the learning environment as simple as possible for training purposes.
-
-DefeatRoaches96 - 
-This variation, DefeatRoaches96, is a 96x96 map that is bigger than the original.
-This allows the agent to have a larger input space to choose from which could potentially improve rate of convergence to an optimal solution.
-
-\section{Results}
-We ran several simulations with various choices of hyperparameters, and our
-results are promising. We saw measurable improvement in preformence after just
-a few iterations of the algorithm, which gave us a steep initial learning curve
-before the improvements started to taper off.
+\section{Results} We ran several simulations with various choices of
+hyper-parameters, and our results were promising. We saw measurable improvement
+in performance after just a few iterations of the algorithm, which gave us a
+steep initial learning curve before the improvements started to taper off.
 
@@
 algorithm. Because we keep the top individual from each generation in the gene
-pool, we see that the top individual has palteaued early, and no new individual
-is achieving a higer fitness. We see that the average is countinually decreasing
-though, which will increase the chances that an individual will be bred that has
-a higher fitness than the individual that is dominating the maximum fitness.
+pool, we see that the top individual has plateaued early, and no new individual
+is achieving a higher fitness. We see that the average is continually increasing
+though, which will increase the chances that an individual will eventually be
+bred that has a higher fitness than the individual that is dominating the
+maximum fitness.
 
 We can also see that the average sometimes plateaus as well. This is due to a
-homonginization of the gene pool, where the children individuals are very
-similar to their parents. This is a case of finding a local maximum. Since we
-favored an exploitive algorithm, this is to be expected, however we see that the
-population can still achieve higher fitness meausures which means that the
-algorithm still retains some amount of exploratory behavior.
+homogenization of the gene pool, where the children individuals are very
+similar to their parents. Because of low mutation rates in this trial, we got
+``unlucky'' and were not creating enough mutations to be exploratory, this is a
+case of finding a local maximum. Since we favored an exploitive algorithm, this
+is to be expected, however we see that the population can still achieve higher
+fitness measures which means that the algorithm still retains some amount of
+exploratory behavior.
 
@@
 regularly. However, despite having more generations, trial two did not reach the
-same average fitness as trial one did. One possible reason for this is since our
-mutation rates are higher, our algorithm favored a more exploratory approach,
-which lead to a more consistant incresase in fitness, but did not reach as high
-of an average fitness.
+same average fitness or top fitness as trial one did. One possible reason for
+this behavior is that our mutation rates are higher, so our algorithm favored a
+more exploratory approach, leading to a more consistent increase in
+fitness. This exploratory behavior also accounts for the top fitness being lower
+than in trial one.
 
-%% TODO: Talk about choice of hyper paramters, population size, episode count
+We did not have enough time to run extensive testing over a representative
+sample of hyper parameters, and expect to get better convergence by tweaking the
+parameters some more. We chose the parameters we did so that we could see
+convergence in a timely manner. Since our simulation times were the bottleneck
+in our pipeline, we choose a relatively small population size to run through our
+algorithm. We think that using higher population sizes would also lead to better
+convergence, as the networks would be put through more and more simulations,
+generating a more and more accurate picture of their fitness.
+
+%% TODO: Talk about choice of hyper parameters, population size, episode count
 %% mutation constants, eta in fitness function
@@
 
-\subsection{Improvements}
-\blindtext
-\blindtext
+\subsection{Improvements}\label{sec:improvements}
+% live back propagation
+One of the for-most improvements we could do is to train the neural networks
+during the simulation as well as training them through evolution. In addition
+to improving the convergence, this strategy would mimic nature even more.
+Organisms learn during their lifetimes, and there is speculation that the knowledge obtained by an individual during its lifetime can be passed down through the
+genome.
+
+We could accomplish this by using backpropogation during the simulations, and considering good measures of progress that we could calculate during the game.
+As an example, we could compute the Euclidean distance between the agent and the adversary and make decisions that minimize this value.
+The drawback to this method is that it may encourage sub-optimal behavior by biasing the agent towards known behavior.
+It is common for human experimenters to expect sub-optimal behavior because optimal behavior is so far unknown \cite{lehman}.
 
@@
 This project can certainly be extended in many ways, for example, we could
-extend the output vectors to include offinsive manuvering by our agent, and
-train the agent to be agressive by considering the number of adversaries killed
+extend the output vectors to include offensive maneuvering by our agent, and
+train the agent to be aggressive by considering the number of adversaries killed
 by our agent during the simulation.
@@
 Another possibility that we did not explore, was the effect of removing neurons
-from the network. Neural networks can often be plauged by overfitting input
+from the network. Neural networks can often be plagued by overfitting input
 data, and by removing neurons we hypothesize that any overfitting we encounter
@@
 removing it from the network. If the neuron in question was indeed problematic,
-we would see it countinue to evolve very low weights, essentially pruning the
+we would see it continue to evolve very low weights, essentially pruning the
 neuron and rendering it vestigial.
@@
 
+\bibitem{pragprog}
+  Andrew Hunt, David Thomas.
+  \textit{The Pragmatic Programmer: From Journeyman to Master}
+  Addison-Wesley (Reading, Mass., 2000)
+
 \end{thebibliography}

### 2018-06-04 doc/run_awai.tex

@@
 simulation controller is built on top of the PySCII interface, and simulates
-a ``game'' between our agent and a group of adversaries, which are controlled by
-the in-game AI that ships with the game.
+a ``game'' with our agent and a group of adversaries, which are controlled by
+the in game AI that come with the game.
 
@@
 \subsection{Controller}
+
+The Pysc2 interface allowed us to create an agent to interact with the Star
+Craft II game.  The agent class contains a \texttt{step()} function that is called
+several times per second of game time. This meant that all of the actions the
+agent took had to be in this step function, and thus this is where we would
+integrate our neural net.  We decided to use the mini game map DefeatRoaches,
+which is included with Pysc2 as a base for our simulation environment. By
+heavily relying on a preexisting map and mini game, we could focus more of our
+time and resources on developing the genetic algorithm and neural network
+implementation.
+
+To make our controller implementation as modular as possible, we decided to
+provide a dictionary of inputs and receive a dictionary of outputs. The inputs
+would be the x and y coordinates of the agent’s units, as well as the x and y
+coordinates of the enemy units. The inputs had to be created and assembled into
+a dictionary in the step function.  In order to normalize the inputs, we divided
+the coordinate value by the maximum map size in the corresponding axis.  To
+obtain the values for the inputs, we created a series of methods for our agent
+class.  These methods include \texttt{getCurrentLocation()} and
+\texttt{getCurrentEnemeyLocation}().  We also created methods to move the friendly units,
+including \texttt{moveToLocation()} and setTargetDestination.  These functions could be
+used together to set a target destination coordinate and cause the friendly
+units to move toward that target.
+
+One problem we encountered that we had to correct was determining when units
+arrived at their target destination.  We solved this by comparing the target
+destination and the units current location every time \texttt{step()} is called.  Because
+the DefeatRoaches mini game uses multiple units (both friendly and hostile), the
+unit location coordinates are averages of all selected units locations.  This
+means that often units will never actually arrive at their precise destination,
+just get very close to it.  To mitigate this issue, we decided to consider units
+within a certain radius of their target destination as at their destination.
+Once this condition is met, the \texttt{moveToLocation()} function returns a status of
+arrived at target.
+
+Whenever \texttt{step()} is called by Pysc2, the first thing we do is check the
+observable action space. If friendly unit movement is not an option, this means
+the unit(s) are not selected.  If units are not selected, this means it is the
+first time the \texttt{step()} function is being called in the simulation episode.  In
+this case, in addition to selecting the units, we also reset various counters
+(such as steps survived).  During the first call of \texttt{step()}, we also unpickle a
+neural network file.
+
+In the case where \texttt{step()} is called and units are already selected, the
+\texttt{moveToLocation()} function is called.  Initially, we set the target destination
+to the enemy location.  We have our units move directly to the hostile units in
+order to engage them in combat, and this initiate a pursuit from the enemy.  We
+felt this would be the most direct way to initiate a pursuit.  We investigated
+running two Pysc2 agents against each other, however Pysc2 currently only
+support the simulation of one custom agent at a time. Once the friendly units
+arrive at their destination, the \texttt{moveToLocation()} function returns an arrived at
+target status to the calling \texttt{step()} function.
+
+When the \texttt{step()} function receives an arrived at target status, a new target
+destination is set.  To generate a new target destination, the relevant aspects
+of the current state (current location of friendly and enemy units) are captured
+and fed into the neural network.  The neural network returns a dictionary
+containing the action space and a value between -1 and 1 for each action label
+key.  The action with the highest number will be chosen, and then taken.
+
+Our action space is movements that the friendly units can take.  We decided to
+quantize the action space into discrete movements, as opposed to a continuous
+x-y coordinate plane. By only allowing our units to move in the 4 cardinal
+directions, diagonals, or stay put, we hoped our first implementation of the
+neural network and genetic algorithm would have an easier time converging.  In
+the future, we could expand the output action space to a continuous x and y
+coordinate plane, as opposed to discrete directions. Although a larger action
+space can often require more iterations of the genetic algorithm to converge, as
+the problem complexity is much higher.
+
+To evaluate the fitness of the neural network, we count the number of steps that
+the units survived for.  One problem we encountered and solved, was the initial
+series of steps taken before engaging the enemy units. For example, the
+DefeatRoaches map spawns the enemy and friendly units on opposite sides of the
+map, in a random selection of several predefined start locations.  However, the
+distance between these start locations can vary between simulation episode.
+This means our neural network fitness (number of steps survived) could be
+slightly affected by an irrelevant condition.  To mitigate this issue, we
+subtract the number of steps it takes until the friendly units reach their first
+destination (the location of the enemy units) and start their engagement. By
+only counting steps after enemy engagement has begun, our fitness function is
+more reflective of survival time.
+
+In order for our evolutionary genetic algorithm to evolve, the end fitness of a
+network must be returned from the Pysc2 simulation.  Our simulation runs a
+configurable (up to 55) episodes per individual (neural network) and averages
+the fitness for all of those simulation episodes. We played with the number of
+simulation episodes per simulation test, exploring the tradeoff between
+simulation runtime and precision.  In the future, we could explore modifying our
+fitness function to incorporate more meaning from having multiple simulation
+episodes.  One future idea we had was to see if the group of individual survival
+times were skewed above or below the mean.
+
+The average fitness for the individual is then saved in a pickle file, along
+with the fitness of the top (most fit) individual.  Because the Pysc2 simulation
+is invoked from the genetic algorithm code, the algorithm can then unpickle the
+file.  After all individuals in a generation are tested and their fitnesses
+collected, the genetic algorithm can select the most fit individuals (neural
+networks) for breeding.
 

### 2018-06-04 doc/run_awai.tex
added sections

--- 2018-06-04 doc/run_awai.tex
@@
 \usepackage{float}
-\usepackage[margin=.75in]{geometry}
+\usepackage[margin=1.25in]{geometry}
 \usepackage{graphicx}
@@
 
-\author{Kyle R. Chickering \\ krchicke@ucdavis.edu
-  \and Alex R. Mirov \\ armirov@ucdavis.edu
-  \and Simon H. Wu \\ simwu@ucdavis.edu
-  \and Xin Jin \\ jin@ucdavis.edu}
+\author{
+    Kyle R. Chickering \\ {\footnotesize\textbf{krchicke@ucdavis.edu}}
+    \and Alex R. Mirov \\ {\footnotesize\textbf{armirov@ucdavis.edu}}
+    \and Simon H. Wu \\   {\footnotesize\textbf{simwu@ucdavis.edu}}
+    \and Xin Jin \\       {\footnotesize\textbf{jin@ucdavis.edu}}
+}
 
@@
 
-\begin{multicols}{2}
+%\begin{multicols}{2}
 \section{Introduction}
@@
 
-\begin{figure}[H]
+\begin{figure}[h]\label{fig:flow chart}
 \centering
-\includegraphics[width=0.45\textwidth]{chart}\label{fig:flow chart}
+\includegraphics[width=0.45\textwidth]{chart}
 \begin{changemargin}{0.15in}{0.15in}
@@
 We chose this function for a number of reasons, the least of which is its
-popularity as an activation function in neural networks. Because of the shape of
-the sigmoid function, inputs to subsequent neurons stay in the nice range of
-$[-1, 1]$, which has the effect of normalizing the network. Since we are not
-using backpropogation, the nice properties of the sigmoid function were
-unimportant to us. However, if we were to do live backpropogation during
+popularity as an activation function in neural networks \cite{russell}. Because
+of the shape of the sigmoid function, inputs to subsequent neurons stay in the
+nice range of $[-1, 1]$, which has the effect of normalizing the network. Since
+we are not using backpropogation, the nice properties of the sigmoid function
+were unimportant to us. However, if we were to do live backpropogation during
 simulations (See section \ref{sec:improvements}), the sigmoid function would be
@@
 \subsection{Controller}
+The controller was the interface we used to communicate with StarCraftII and run
+the simulations need to commute the fitness function. Our controller has a
+function to step the game simulation, which is called several times per
+second. We implemented all of our agent's logic, consisting of feeding forward
+the inputs through our neural network, inside of this step function.
+
+We spent some time deciding what inputs to send to our neural network during the
+simulation. Originally, we sent absolute coordinates, but we found that these
+coordinates made the network into essentially a binary decision
+network. Instead, we settled on giving the neural networks relative coordinates
+produced much better results. We also give the neural network the relative
+coordinates of the enemy units. The controller queries the PySC2 interface for
+these values every time we take a step in the simulation, and the values are
+pulled out of StarCraftII itself.
+
+% The Pysc2 interface allowed us to create an agent to interact with the Star
+% Craft II game.  The agent class contains a \texttt{step()} function that is
+% called several times per second of game time. This meant that all of the actions
+% the agent took had to be in this step function, and thus this is where we would
+% integrate our neural net.  We decided to use the mini game map DefeatRoaches,
+% which is included with Pysc2 as a base for our simulation environment. By
+% heavily relying on a preexisting map and mini game, we could focus more of our
+% time and resources on developing the genetic algorithm and neural network
+% implementation.
+
+% To make our controller implementation as modular as possible, we decided to
+% provide a dictionary of inputs and receive a dictionary of outputs. The inputs
+% would be the x and y coordinates of the agent’s units, as well as the x and y
+% coordinates of the enemy units. The inputs had to be created and assembled into
+% a dictionary in the step function.  In order to normalize the inputs, we divided
+% the coordinate value by the maximum map size in the corresponding axis.  To
+% obtain the values for the inputs, we created a series of methods for our agent
+% class.  These methods include \texttt{getCurrentLocation()} and
+% \texttt{getCurrentEnemeyLocation}().  We also created methods to move the
+% friendly units, including \texttt{moveToLocation()} and setTargetDestination.
+% These functions could be used together to set a target destination coordinate
+% and cause the friendly units to move toward that target.
+
+One problem we encountered that we had to correct was determining when units
+arrived at their target destination.  We solved this by comparing the target
+destination and the unit's current location every time we step the simulation.
+Because the mini game uses multiple units (both friendly and hostile), the unit
+location coordinates are averaged over all selected units locations.  This means
+that often units will never actually arrive at their precise destination, just
+get very close to it. To mitigate this issue, we decided to consider units
+within a certain radius of their target destination as at their destination.
+Once this condition is met, we could ask the simulation whether or not units had
+reached their destination. Once the units reach their destination, we can choose
+a new destination for them to move to.
+
+% Whenever \texttt{step()} is called by Pysc2, the first thing we do is check the
+% observable action space. If friendly unit movement is not an option, this means
+% the unit(s) are not selected.  If units are not selected, this means it is the
+% first time the \texttt{step()} function is being called in the simulation
+% episode.  In this case, in addition to selecting the units, we also reset
+% various counters (such as steps survived).  During the first call of
+% \texttt{step()}, we also unpickle a neural network file.
+
+The controller also handles the interpretation of the data from the PySC2 interface, and acts as a ``human'' playing the game.
+The controller is responsible for selecting movements, making decisions, and executing those decisions.
+
+% In the case where \texttt{step()} is called and units are already selected, the
+% \texttt{moveToLocation()} function is called.  Initially, we set the target
+% destination to the enemy location.  We have our units move directly to the
+% hostile units in order to engage them in combat, and this initiate a pursuit
+% from the enemy.  We felt this would be the most direct way to initiate a
+% pursuit.  We investigated running two Pysc2 agents against each other, however
+% Pysc2 currently only support the simulation of one custom agent at a time. Once
+% the friendly units arrive at their destination, the \texttt{moveToLocation()}
+% function returns an arrived at target status to the calling \texttt{step()}
+% function.
+
+When the controller receives an arrived at target status, a new
+target destination is set.  To generate a new target destination, the relevant
+aspects of the current state (current location of friendly and enemy units) are
+captured and fed into the neural network.  The neural network returns a
+vector of the action space and a value between -1 and 1 for each
+action label key. The action with the highest number will be chosen, and then
+taken.
+
+Our action space is movements that the friendly units can take. We decided to
+quantize the action space into discrete movements, as opposed to a continuous
+$x$-$y$ coordinate plane. By only allowing our units to move in the 4 cardinal
+directions, diagonals, or stay put, we hoped our first implementation of the
+neural network and genetic algorithm would have an easier time converging.
+
+%% Fix
+To evaluate the fitness of the neural network, we count the number of steps that
+the units survived for.  One problem we encountered and solved, was the initial
+series of steps taken before engaging the enemy units. For example, the
+DefeatRoaches map spawns the enemy and friendly units on opposite sides of the
+map, in a random selection of several predefined start locations.  However, the
+distance between these start locations can vary between simulation episode.
+This means our neural network fitness (number of steps survived) could be
+slightly affected by an irrelevant condition.  To mitigate this issue, we
+subtract the number of steps it takes until the friendly units reach their first
+destination (the location of the enemy units) and start their engagement. By
+only counting steps after enemy engagement has begun, our fitness function is
+more reflective of survival time.
+
+% In order for our evolutionary genetic algorithm to evolve, the end fitness of a
+% network must be returned from the Pysc2 simulation.  Our simulation runs a
+% configurable (up to 55) episodes per individual (neural network) and averages
+% the fitness for all of those simulation episodes. We played with the number of
+% simulation episodes per simulation test, exploring the tradeoff between
+% simulation runtime and precision.  In the future, we could explore modifying our
+% fitness function to incorporate more meaning from having multiple simulation
+% episodes.  One future idea we had was to see if the group of individual survival
+% times were skewed above or below the mean.
+
+In addition, the controller is responsible for calculating $\mathcal{S}(e)$ for
+each episode (simulation) that it runs, and returning the value to our genetic
+algorithm. This allows us to make modifications to our fitness function without
+having to make changes to the genetic algorithm. This also gives us the
+flexibility to extend the capabilities of the controller to run different style
+simulations in the future. For example we could run the agent through multiple
+types of maps, and compute the average fitness between multiple different
+episodes and maps.
+
+The average fitness for the individual is then saved in a Python pickle file,
+along with the fitness of the top (most fit) individual.  Because the Pysc2
+simulation is invoked from the genetic algorithm code, the algorithm can then
+unpickle the file.  After all individuals in a generation are tested and their
+fitnesses collected, the genetic algorithm can select the most fit individuals
+(neural networks) for breeding.
 
@@
 
-\begin{figure}[H]\label{fig:fitness graph}
+\begin{figure}[h]\label{fig:fitness graph}
 \centering
@@
 
-%% TODO: Talk about choice of hyper parameters, population size, episode count
-%% mutation constants, eta in fitness function
-
 \section{Conclusion}
 Different selection types?
-\blindtext
 
 \subsection{Improvements}\label{sec:improvements}
-% live back propagation
-One of the for-most improvements we could do is to train the neural networks
+One of the foremost improvements we could do is to train the neural networks
 during the simulation as well as training them through evolution. In addition
@@
 by our agent during the simulation.
+
+In the future, we could expand the output action space to a continuous $x$ and
+$y$ coordinate plane, as opposed to discrete directions. Although a larger
+action space can often require more iterations of the genetic algorithm to
+converge, the benefits would be more nuanced agent behavior, and more realistic
+agent movements. As it stands now, our agent moves in what may be considered a
+very unnatural way, and while this does not affect the agent's ability to
+preform, it could be seen as a disadvantage in certain situations where the fact
+that the game is being played by an AI may be undesirable.
+
+We have also talked about putting more attention towards our fitness function.
+An interesting possibility is to take into account the number of episodes in $\bm{\mathcal{E}}$, and rewarding agents that have more simulations.
+We can then associate individuals with their simulation meta-data, and save time running simulations.
+We can start with a low number of simulations per individual, and as the generations progress, increase the number of simulations per individual.
+This would have the effect of starting with relatively low fitness, but increasing the ``accuracy'' of the fitness over time.
 
@@
 
-\end{multicols}
+%\end{multicols}
 
@@
 \begin{thebibliography}{9}
-\bibitem{NEAT} 
-  Kenneth O. Stanley, Risto Miikkulainen.
-  \textit{Evolving Neural Networks through Augmenting Topologies}. 
-  The MIT Press Journals.
 
 \bibitem{dejong}
-  Kenneth A. De Jong. \textit{Evolutionary Computation: A Unified Approach}.
+  Kenneth A. De Jong.
+  \textit{Evolutionary Computation: A Unified Approach}.
   MIT Press, Feb 3, 2006.
-
-\bibitem{russell}
-  Stuart Russell, Peter Norvig.
-  \textit{Artificial Intelligence: A Modern Approach}.
-  Upper Saddle River (New Jersey, 1995).
-
-\bibitem{lehman}
-  Joel Lehman et al.
-  \textit{The Surprising Creativity of Digital Evolution: A Collection of
-    Anecdotes from the Evolutionary Computation and Artificial Life Research
-    Communities}.
-  \textbf{https://arxiv.org/pdf/1803.03453.pdf}.
 
@@
 
+\bibitem{lehman}
+  Joel Lehman et al.
+  \textit{The Surprising Creativity of Digital Evolution: A Collection of
+    Anecdotes from the Evolutionary Computation and Artificial Life Research
+    Communities}.
+  \textbf{https://arxiv.org/pdf/1803.03453.pdf}.
+
+\bibitem{russell}
+  Stuart Russell, Peter Norvig.
+  \textit{Artificial Intelligence: A Modern Approach}.
+  Upper Saddle River (New Jersey, 1995).
+
+\bibitem{NEAT} 
+  Kenneth O. Stanley, Risto Miikkulainen.
+  \textit{Evolving Neural Networks through Augmenting Topologies}. 
+  The MIT Press Journals.
+
 \end{thebibliography}

### 2018-06-04 doc/run_awai.tex
short conclusion

@@
 \section{Conclusion}
-Different selection types?
+Based on the convergence we got during our trials, our method of training a
+neural network to run away is a feasable approach to the problem. Although we
+arguably did not achieve ``optimal'' results, given more time we are confident
+that we could significantly improve the convergence of our algorithm.
+
+We have demonstrated that evolving neural networks using a genetic algorithm is
+a good approach to teaching agents to accomplish goals in countinuous or near
+countinuous state spaces.
 

### 2018-06-04 util/generate_graphs.py
add data graphing utility

@@ -1,0 +1,26 @@
+import csv
+import matplotlib.pyplot as plt
+import numpy as np
+import sys
+
+if __name__ == "__main__":
+    data_avg = []
+    data_top = []
+    
+    with open(sys.argv[1], 'r') as csvfile:
+        reader = csv.reader(csvfile, delimiter=',')
+        
+        for row in reader:
+            data_avg.append(500 - float(row[1].strip()))
+            data_top.append(500 - float(row[2].strip()))
+
+        csvfile.close()
+
+    plt.plot(data_avg)
+    plt.plot(data_top)
+    plt.xlabel('Generation')
+    plt.ylabel('Average Fitness\nTop Fitness')
+    plt.ylim(0, 300);
+    plt.show()
+
+

### 2018-06-06 doc/run_awai.tex
fixed image formatting
added abstract

@@
 \begin{abstract}
-  \textbf{Abstract}: \blindtext
+  \textbf{Abstract}:
+
+  This project explores the integration of a neural network and an evolutionary
+  genetic algorithm with the Star Craft II API.The goal of this project was to
+  successfully control a StarCraft II agent via a neural network, and maximize
+  its fitness through a genetic evolutionary algorithm. The agent uses the PySC2
+  API to connect with StarCraft II, and a neural network to control its
+  actions.The fitness function is defined as the survival duration for an agent
+  in a sandbox environment with hostile units, where surviving for longer is
+  better.The action space of the agent was limited to movement in order to
+  achieve a realistic scope for the project.The fitness of an individual is
+  evaluated by running multiple simulation episodes and averaging the
+  performance (survival time) for each episode.The neural network uses an
+  explicit topologically unrestricted implementation.This implementation
+  provides high degrees of freedom while minimizing the network size.The genetic
+  algorithm uses tournament selection to select the most fit neural network
+  (individuals) to breed.Each successive generation of neural networks is
+  slightly better at surviving in the hostile sandbox environment.After 50
+  iterative generations, our resulting neural network had increased the average
+  survival time of an agent by 288%.
+
   \\~\\
@@
 because we were curious if the flexible topologies would lead to a more
-efficient neural network. We were inspired by the NEAT algorithm \cite{NEAT}, which has
-shown promise in other video game domains. However we choose a simpler
+efficient neural network. We were inspired by the NEAT algorithm \cite{NEAT},
+which has shown promise in other video game domains. However we choose a simpler
 topologically explicit implementation to simplify and accelerate the development
-of our
-algorithm. We ultimately settled on an acyclic graph representation, which
-allowed us the freedom to add nodes and allowed simple manipulation of
+of our algorithm. We ultimately settled on an acyclic graph representation,
+which allowed us the freedom to add nodes and allowed simple manipulation of
 connection weights.
@@
 
-% The Pysc2 interface allowed us to create an agent to interact with the Star
-% Craft II game.  The agent class contains a \texttt{step()} function that is
-% called several times per second of game time. This meant that all of the actions
-% the agent took had to be in this step function, and thus this is where we would
-% integrate our neural net.  We decided to use the mini game map DefeatRoaches,
-% which is included with Pysc2 as a base for our simulation environment. By
-% heavily relying on a preexisting map and mini game, we could focus more of our
-% time and resources on developing the genetic algorithm and neural network
-% implementation.
-
-% To make our controller implementation as modular as possible, we decided to
-% provide a dictionary of inputs and receive a dictionary of outputs. The inputs
-% would be the x and y coordinates of the agent’s units, as well as the x and y
-% coordinates of the enemy units. The inputs had to be created and assembled into
-% a dictionary in the step function.  In order to normalize the inputs, we divided
-% the coordinate value by the maximum map size in the corresponding axis.  To
-% obtain the values for the inputs, we created a series of methods for our agent
-% class.  These methods include \texttt{getCurrentLocation()} and
-% \texttt{getCurrentEnemeyLocation}().  We also created methods to move the
-% friendly units, including \texttt{moveToLocation()} and setTargetDestination.
-% These functions could be used together to set a target destination coordinate
-% and cause the friendly units to move toward that target.
-
 One problem we encountered that we had to correct was determining when units
@@
 
-% Whenever \texttt{step()} is called by Pysc2, the first thing we do is check the
-% observable action space. If friendly unit movement is not an option, this means
-% the unit(s) are not selected.  If units are not selected, this means it is the
-% first time the \texttt{step()} function is being called in the simulation
-% episode.  In this case, in addition to selecting the units, we also reset
-% various counters (such as steps survived).  During the first call of
-% \texttt{step()}, we also unpickle a neural network file.
-
-The controller also handles the interpretation of the data from the PySC2 interface, and acts as a ``human'' playing the game.
-The controller is responsible for selecting movements, making decisions, and executing those decisions.
-
-% In the case where \texttt{step()} is called and units are already selected, the
-% \texttt{moveToLocation()} function is called.  Initially, we set the target
-% destination to the enemy location.  We have our units move directly to the
-% hostile units in order to engage them in combat, and this initiate a pursuit
-% from the enemy.  We felt this would be the most direct way to initiate a
-% pursuit.  We investigated running two Pysc2 agents against each other, however
-% Pysc2 currently only support the simulation of one custom agent at a time. Once
-% the friendly units arrive at their destination, the \texttt{moveToLocation()}
-% function returns an arrived at target status to the calling \texttt{step()}
-% function.
+The controller also handles the interpretation of the data from the PySC2
+interface, and acts as a ``human'' playing the game. The controller is
+responsible for selecting movements, making decisions, and executing those
+decisions.
 
@@
 
-%% Fix
 To evaluate the fitness of the neural network, we count the number of steps that
@@
 more reflective of survival time.
-
-% In order for our evolutionary genetic algorithm to evolve, the end fitness of a
-% network must be returned from the Pysc2 simulation.  Our simulation runs a
-% configurable (up to 55) episodes per individual (neural network) and averages
-% the fitness for all of those simulation episodes. We played with the number of
-% simulation episodes per simulation test, exploring the tradeoff between
-% simulation runtime and precision.  In the future, we could explore modifying our
-% fitness function to incorporate more meaning from having multiple simulation
-% episodes.  One future idea we had was to see if the group of individual survival
-% times were skewed above or below the mean.
 
@@
 One of the foremost improvements we could do is to train the neural networks
-during the simulation as well as training them through evolution. In addition
-to improving the convergence, this strategy would mimic nature even more.
-Organisms learn during their lifetimes, and there is speculation that the knowledge obtained by an individual during its lifetime can be passed down through the
-genome.
+during the simulation as well as training them through evolution. In addition to
+improving the convergence, this strategy would mimic nature even more.
+Organisms learn during their lifetimes, and there is speculation that the
+knowledge obtained by an individual during its lifetime can be passed down
+through the genome.
 
-We could accomplish this by using backpropogation during the simulations, and considering good measures of progress that we could calculate during the game.
-As an example, we could compute the Euclidean distance between the agent and the adversary and make decisions that minimize this value.
-The drawback to this method is that it may encourage sub-optimal behavior by biasing the agent towards known behavior.
-It is common for human experimenters to expect sub-optimal behavior because optimal behavior is so far unknown \cite{lehman}.
+We could accomplish this by using backpropogation during the simulations, and
+considering good measures of progress that we could calculate during the
+game. As an example, we could compute the Euclidean distance between the agent
+and the adversary and make decisions that minimize this value. The drawback to
+this method is that it may encourage sub-optimal behavior by biasing the agent
+towards known behavior. It is common for human experimenters to expect
+sub-optimal behavior because optimal behavior is so far unknown \cite{lehman}.
 
@@
 
-We have also talked about putting more attention towards our fitness function.
-An interesting possibility is to take into account the number of episodes in $\bm{\mathcal{E}}$, and rewarding agents that have more simulations.
-We can then associate individuals with their simulation meta-data, and save time running simulations.
-We can start with a low number of simulations per individual, and as the generations progress, increase the number of simulations per individual.
-This would have the effect of starting with relatively low fitness, but increasing the ``accuracy'' of the fitness over time.
+We have also talked about putting more attention towards our fitness
+function. An interesting possibility is to take into account the number of
+episodes in $\bm{\mathcal{E}}$, and rewarding agents that have more
+simulations. We can then associate individuals with their simulation meta-data,
+and save time running simulations. We can start with a low number of simulations
+per individual, and as the generations progress, increase the number of
+simulations per individual. This would have the effect of starting with
+relatively low fitness, but increasing the ``accuracy'' of the fitness over
+time.
 

### 2018-06-06 doc/run_awai.tex

@@
   This project explores the integration of a neural network and an evolutionary
-  genetic algorithm with the Star Craft II API.The goal of this project was to
+  genetic algorithm with the Star Craft II API. The goal of this project was to
   successfully control a StarCraft II agent via a neural network, and maximize
@@
   API to connect with StarCraft II, and a neural network to control its
-  actions.The fitness function is defined as the survival duration for an agent
+  actions. The fitness function is defined as the survival duration for an agent
   in a sandbox environment with hostile units, where surviving for longer is
-  better.The action space of the agent was limited to movement in order to
-  achieve a realistic scope for the project.The fitness of an individual is
+  better. The action space of the agent was limited to movement in order to
+  achieve a realistic scope for the project. The fitness of an individual is
   evaluated by running multiple simulation episodes and averaging the
-  performance (survival time) for each episode.The neural network uses an
-  explicit topologically unrestricted implementation.This implementation
-  provides high degrees of freedom while minimizing the network size.The genetic
+  performance (survival time) for each episode. The neural network uses an
+  explicit topologically unrestricted implementation. This implementation
+  provides high degrees of freedom while minimizing the network size. The genetic
   algorithm uses tournament selection to select the most fit neural network
-  (individuals) to breed.Each successive generation of neural networks is
-  slightly better at surviving in the hostile sandbox environment.After 50
+  (individuals) to breed. Each successive generation of neural networks is
+  slightly better at surviving in the hostile sandbox environment. After 50
   iterative generations, our resulting neural network had increased the average
-  survival time of an agent by 288%.
+  survival time of an agent by 288\%.
 
@@
 \centering
-\includegraphics[width=0.45\textwidth]{chart}
+\includegraphics[width=0.75\textwidth]{chart}
 \begin{changemargin}{0.15in}{0.15in}
@@
 \centering
-\includegraphics[width=0.45\textwidth]{fig_1}
+\includegraphics[width=0.75\textwidth]{fig_1}
 \begin{changemargin}{0.15in}{0.15in}
@@
 \centering
-\includegraphics[width=0.45\textwidth]{fig_2}
+\includegraphics[width=0.75\textwidth]{fig_2}
 \begin{changemargin}{0.15in}{0.15in}

### 2018-06-07 nn/nn3.py
Fixed import,Fixed cycle issue,

@@ -29,7 +29,7 @@ class NeuralNetwork(object):
                  learning_constant=1, struct_mut_con_rate=0.5,
                  struct_mut_new_rate=0.5, struct_mut_rm_rate=0.5,
                  n_struct_mut_rate=0.001):
-        """ Configures network parameters for this neural network. 
+        """ Configures network parameters for this neural network.
 
         input_size: The number of inputs to the neural network.
         output_size: The number of outputs from the neural network.
@@ -46,18 +46,18 @@ class NeuralNetwork(object):
           which new connections form in the network. A 0 is for new connections
           never being built, and a 1 guarenting a new connection if a new
           connection is possible.
-        struct_mut_new_rate: A number between 0 and 1 specifing the rate at 
+        struct_mut_new_rate: A number between 0 and 1 specifing the rate at
           which new neurons are added to the network. A 0 adds no new nodes to
           the network, a 1 gaurentees a new node is added at each structural
           mutation.
-        struct_mut_rm_rate: This parameter is not currently implemented. 
+        struct_mut_rm_rate: This parameter is not currently implemented.
           Specifies the rate at which nodes can be spontaniously removed from
-          the network. A number between 0 and 1. 0 is no removal, 1 is a 
+          the network. A number between 0 and 1. 0 is no removal, 1 is a
           gaurenteed removal each structural mutation.
         n_struct_mut_rate:
 
         return: None """
-        
+
         self.g_dict = {}
 
         self.input_size = input_size
@@ -83,7 +83,7 @@ class NeuralNetwork(object):
 
     def picklable(self):
         data = {}
-        
+
         data['g_dict'] = self.g_dict
         data['input_size'] = self.input_size
         data['input_labels'] = self.input_labels
@@ -123,7 +123,7 @@ class NeuralNetwork(object):
         self.struct_mut_rm_rate = data['struct_mut_rm_rate']
         self.n_struct_mut_rate = data['n_struct_mut_rate']
         self.n_struct_mut_minimizer = data['n_struct_mut_minimizer']
-        
+
         self.learning_constant = data['learning_constant']
 
 
@@ -131,7 +131,7 @@ class NeuralNetwork(object):
         """ Prints all of the connections in the neural network.
 
         return: None """
-        
+
         cons = {}
         for node in self.g_dict:
             cons[node] = []
@@ -143,13 +143,13 @@ class NeuralNetwork(object):
         print(cons)
         return
 
-    
+
     def bfs_reorder(self):
[…]
